---
title: "Ross Foundation 2019 Toy Giveaway Data Report"
author: "Sam Parmar"
date: "December 2019"
output:
  html_document: default
  pdf_document: default
---
###### Purpose: Assess Ross Foundation Christmas 2019 Toy Giveaway Survey Data 
###### Significance: Obtain actionable findings from most recent intake data
###### Limitations and Warnings: Possible response bias 

#### First lets load the R packages we need before doing our analysis
```{r message=FALSE, warning=FALSE, error=FALSE, echo=FALSE}
#1)Load packages and set graphs theme
ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}
ipak(c("tidyverse","rio","stringr","ggthemes","dplyr","knitr","kableExtra"))
theme_set(theme_minimal())
```

```{r message=FALSE, warning=FALSE, error=FALSE, echo=FALSE}
#2)Import data
path <- "/Users/Sam/Documents/Programming/R_Prog/2019 toy drive/Ross Foundations Services Referral form - Toy Drive 2019.csv"
rd<-import(path)
```

```{r message=FALSE, warning=FALSE, error=FALSE, echo=FALSE}
#3)Perform data standardization and create a few variables
rd <- rd %>% mutate(`Children age ranges` = `Children age ranges  (check all the apply)`)

rd1<- rd %>% mutate(
  Ethnicity=ifelse(is.na(`Race/Ethnicity`)|str_detect(`Race/Ethnicity`,paste(c("bi-racial","multi-racial","Multiple","Multiracial","White & Puerto Rican"),collapse="|")) , "Missing or Other", `Race/Ethnicity`),NumChiTheor=ifelse(str_detect(`Children age ranges`,"16-19"), `Number of Children in Household`-1, `Number of Children in Household`),
  NumChiTheor=ifelse(is.na(NumChiTheor), str_count(`Children age ranges`,"-"),NumChiTheor), NumChildEvent = `Number of Children with you Today`) 

rd2 <- rd1


rd2 <- rd2 %>% mutate(Child_Infant=ifelse(str_detect(`Children age ranges`,"Infant"), 1,0),
                      Child_1to3=ifelse(str_detect(`Children age ranges`,"1-3"), 1,0),
                      Child_4to7=ifelse(str_detect(`Children age ranges`,"4-7"), 1,0),
                      Child_8to11=ifelse(str_detect(`Children age ranges`,"8-11"), 1,0),
                      Child_12to15=ifelse(str_detect(`Children age ranges`,"12-15"), 1,0),
                      Child_16to19=ifelse(str_detect(`Children age ranges`,"16-19"), 1,0)
)

rd2 <- rd2 %>% mutate(F_Housing=ifelse(str_detect(`Please check all issues that apply to this FAMILY`,"Housing"), 1,0),
                      F_EmploymentFin=ifelse(str_detect(`Please check all issues that apply to this FAMILY`,"Employment/financial"), 1,0),
                      F_BasicNeeds=ifelse(str_detect(`Please check all issues that apply to this FAMILY`,"Basic needs (food, clothing)"), 1,0),
                      F_Legal=ifelse(str_detect(`Please check all issues that apply to this FAMILY`,
                                                "Legal"), 1,0),
                      F_HealthIns=ifelse(str_detect(`Please check all issues that apply to this FAMILY`,"Health Insurance"), 1,0),
                      F_HealthMH=ifelse(str_detect(`Please check all issues that apply to this FAMILY`,"Health/mental health"), 1,0),
                      F_fsrissues=ifelse(str_detect(`Please check all issues that apply to this FAMILY`,"Parent/family/student relationship issues"), 1,0)
)

rd2 <- rd2 %>% mutate(C_Academics=ifelse(str_detect(`Please check all issues that apply to CHILDREN`,"Academics"), 1,0),
                      C_Attendance=ifelse(str_detect(`Please check all issues that apply to CHILDREN`,"Attendance"), 1,0),
                      C_EmotBehav=ifelse(str_detect(`Please check all issues that apply to CHILDREN`,"Emotional/Behavioral support"), 1,0),
                      C_MedIss=ifelse(str_detect(`Please check all issues that apply to CHILDREN`,
                                                 "Medical issues"), 1,0),
                      C_Discipline=ifelse(str_detect(`Please check all issues that apply to CHILDREN`,"Discipline"), 1,0)
)

rd2 <- rd2 %>% mutate(H_Friends=ifelse(str_detect(`How did you hear about this event?`,"Friends"), 1,0),
                      H_Family=ifelse(str_detect(`How did you hear about this event?`,"Family member"), 1,0),
                      H_Website=ifelse(str_detect(`How did you hear about this event?`,"Our website"), 1,0),
                      H_News=ifelse(str_detect(`How did you hear about this event?`,
                                                 "News"), 1,0),
                      H_Radio=ifelse(str_detect(`How did you hear about this event?`,"Radio"), 1,0),
                      H_SocialMedia=ifelse(str_detect(`How did you hear about this event?`,"Social Media"), 1,0)
)


labs <- c(paste(seq(0, 95, by = 5), seq(0 + 5 - 1, 100 - 1, by = 5),
                sep = "-"), paste(100, "+", sep = ""))
rd2$Age <- as.numeric(rd2$Age)
rd2$AgeGroup <- cut(rd2$Age, breaks = c(seq(0, 100, by = 5), Inf), labels = labs, right = FALSE)

#remove unnecessary data frames from env
rm(rd,rd1)
```

#### 1)Estimated number of children that attended our Toy Drive:
```{r message=FALSE, warning=FALSE, error=FALSE, echo=FALSE}
rd2 %>% select(NumChildEvent) %>% 
  summarise_all(funs(mean, max, min,sum), na.rm = TRUE) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```
* At least 414 children attended our event and had a parent who completed intake forms
* Note that there were 6 parents that did not fill in how many children were with them


#### 2) Parent demographics info:
```{r message=FALSE, warning=FALSE, error=FALSE, echo=FALSE}



rd2 %>% summarize(CountParents=n()) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F)


#Breakdown of Parents that attended by Ethnicity:
rd2 %>% group_by(Ethnicity) %>% summarize(CountParents=n()) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F)


#Breakdown of Parents by Age and Race/Ethnicities
rd2 %>% group_by(`AgeGroup`) %>% summarize(Total=n()) %>% merge(., (rd2 %>% group_by(`AgeGroup`, Ethnicity) %>% summarize(CountNames=n()) %>% spread(Ethnicity, CountNames)),by="AgeGroup") %>% arrange(AgeGroup) %>%
kable() %>%
kable_styling(bootstrap_options = "striped", full_width = F)
```
* Roughly 179 parents attended our 2019 Toy Drive and completed intake forms
    * When looking by race/ethnicity, the majority were Black.
    * When looking by age group, the majority were 25-29.

#### 3) Reported family issues/needs:
```{r message=FALSE, warning=FALSE, error=FALSE, echo=FALSE} 
#Family Issues/Needs from Parents that attended our Toy Drive (responses not mutually exclusive)
rd2 %>% select(starts_with("F_")) %>% 
  summarise_all(funs(sum), na.rm = TRUE) %>%  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F)

#Children Issues/Needs from Parents that attended our Toy Drive (responses not mutually exclusive)
rd2 %>% select(starts_with("C_")) %>% 
  summarise_all(funs(sum), na.rm = TRUE) %>%  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```
* For family issues (upper table): 
    * We see the most parents reported housing issues (F_housing), health insurance issues (F_HealthIns), and then mental health issues (F_HealthMH). 
    * We also asked about issues associated with employement/financial (F_EmployementFin), basic needs like food, clothing (F_BasicNeeds), Legal (F_Legal), and Parent/family/student relationship (F_fsrissues)

* For children issues (lower table):
    * We see the most parents reported emotional behavioral issues (C_EmotBehav), then discipline issues (C_Discipline), then academic issues (C_Academic) or medical issues (C_MedIss). We also asked about attendance issues (C_Attendance) This highlights the importance of student support services at our events. 
* Note the individuals were allowed to select one or more of the issue categories.

#### 4) Requesting contact with assistance for any issues listed on this form
```{r message=FALSE, warning=FALSE, error=FALSE, echo=FALSE} 
#Would you like the Ross Foundation to contact you to help with assistance 
#for any issues listed on this form?
rd2 %>%   count(`Would you like the Ross Foundation to contact you to help with assistance for any issues listed on this form?`) %>% kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```
* 127 parents requested to be contacted by the Ross Foundation for assistance with children and/or family issues


#### 5) How did they hear about the event?
```{r message=FALSE, warning=FALSE, error=FALSE, echo=FALSE} 
#Children Issues/Needs from Parents that attended our Toy Drive (responses not mutually exclusive)
rd2 %>% select(starts_with("H_")) %>% 
  summarise_all(funs(sum), na.rm = TRUE) %>%  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```
* Most people reported hearing about our event via Social Media or Friends.
* Note that individuals were allows to select one or more of the categories on how they heard about the event.

